\framework: Morphology-based Entity and Relational Entity Extraction Framework for Arabic

### 2018-04-20 tex/merf/TALMajor/tool.tex
done till 3.19

\input{MERFOverview}

The \framework framework is illustrated in the flow diagram of
Figure~\ref{f:overview}.
The Arabic text and the reference tag chunks are the primary inputs to \framework. 
Solutions, morphology-based Boolean formulae, tags, 
morphology-based regular expressions, 
tag chunks, relation and action definitions, and data structures expressing entities 
and relations are input and output data to processes. 
The morphological analyzer (Sarf), $Syn^k$ detector, 
GUI for Boolean formulae definition, 
visualization annotator, GUI for regular expression and action definition, 
Boolean formula simulator, regular expression simulator, 
relation extraction and action execution, and difference and statistical analyzer 
are processes.

@@

\major{
Let ${\cal O} = \{ \mathit{isA}, \mathit{contains}\}$ be the set of atomic term 
predicates, where $\mathit{isA}$ and $\mathit{contains}$ denote exact match and containment, respectively.
Also, let ${\cal F} = \{ \pp, \ss, \xx, \PP, \GG, \AC\}$ be the 
set of morphological features where each entry is in turn a set of values for that feature.
Given a word $w$, a morphological feature set $A\in{\cal F}$,}
a user defined constant feature value $CF\in A$,
and an integer $k, 1\le k\le 7$, 
the following are morphology-based atomic terms (MAT), {\em terms} for short.
\begin{itemize}
  \item $a(w):= \exists m \in M(w).~m=\langle p,s,x,P,G,C\rangle. r \circ CF$
where $\circ \in {\cal O}$, $r \in \{p,s,x,P,G,C\}$, and $r\in A$.
Informally, a solution vector of $w$ exists with
a feature containing or exactly matching \major{the user-chosen feature value} $CF$.
\item $a(w) := w \in Syn^k(CF), CF \in \ss$.
  Informally, this checks if $w$ is an extended synonym of a stem $CF$. 
\major{
  We limit $k$ to a maximum of $7$ since we practically noticed that 
  (1) values above $7$ introduce significant semantic noise and
  (2) the computation is expensive without a bound. 
}
\end{itemize}

@@

Consider the task we discussed in the introduction (Figure~\ref{fig:intromotiv}) 
and recall that we are interested in identifying names of people, names of places, relative positions, and numerical terms.
The table in Figure~\ref{fig:taskMBF} presents the defined formulae.
The user denotes the ``name of person'' entities with formula $N$ 
which requires the {\em category} feature in the morphological solution of a word to be {\tt Name\_of \_Person}.
The entities $n_1$, $n_2$, and $n_3$ are matches of the formula $N$ in the text.
Similarly, the user specifies formula $P$ to denote ``name of place'' entities. 
The user specifies formula $R$ to denote ``relative position'' entities, 
and defines it as a disjunction of MAT terms that check for 
solutions matching stems such as \RL{qrb} (near) and \RL{fy} (in).
%requires the stem feature of a word to belong to a selected list of stems containing \RL{fy} and \RL{qrb}. a
Similarly, $U$ denotes numerical terms and is a disjunction of constraints 
requiring the stem feature to belong to a set of stems such as
\RL{'wl}(first), \RL{_tAny}(second), \dots \RL{`A^sr}(tenth).

@@

Next, we define a morphology-based regular expression (MRE) as follows.
\begin{itemize}
\item $m$ is an MRE where $m$ is an MBF.
\item $fg$ is an MRE denoting a concatenation operation
  where $f$ and $g$ are both MRE expressions. This is 
  satisfied by a match of $f$ followed by a match of $g$. 
\item $f*,f+,f$\textasciicircum$x,$ and $f?$ are MRE expressions where $f$ is an MRE,
  and are satisfied by zero or more matches, one of more matches, up to $x$ matches, 
  and zero or only one match of $f$, respectively.
\item $f\& g,$ (conjunction) and $f|g$ (disjunction) are  MRE expressions where 
  $f$ and $g$ are MRE expressions, and are satisfied by a 
  a match of both $f$ and $g$, 
  and a match of either $f$ or $g$, respectively. 
\end{itemize}
We denote by $\ldbrack f\rdbrack$ the set of matches of an MRE $f$.

Back to the example in Figure~\ref{fig:intromotiv}.
We use the formulae defined in Figure~\ref{fig:taskMBF} to construct an MRE such as
$(P|N)\!+O?~R~O^\wedge\!2~(P|N|U)+$ where 
$|,+,?,$ and $^\wedge k$ denote disjunction, one or more, zero or one, and
up to $k$ matches, respectively.
The expression specifies a sequence of places or names of persons, 
optionally followed by a null word, 
followed by one relative position, followed by up to two possible null words, 
followed by one or more match of name of place, name of person, or numerical term.
$O?$ and $O^\wedge 2$ are used in the expression to allow for flexible matches.
%Note that the user reaches the satisfying matches by experimenting with the visualizer and the expression editor which do not require knowledge and expertise in regular expressions.

@@

The match trees in Figure~\ref{fig:taskMRE} illustrate two matches of the expression computed by \framework.
The first match tree refers to the text 
\RL{brj _hlyfT bAlqrb mn AltqA.t` Al-'wl}(Khalifa Tower next to the first intersection).
The second match tree refers to the text 
\RL{dby mwl `l_A mqrbT mn h_dA Almbn_A}(Dubai Mall is located near this building).
%The leaf nodes of the trees are entities and the edges and internal nodes are text, 
%morphology-based, 
%and word distance based relational entities. 
\major{
The leaf nodes of the match trees are matches to formulae and the internal nodes represent roots to sub-expression matches.
For instance, \RL{brj _hlyfT} in match 1 tree corresponds to the sub-expression $(P|N)+$.
}

\subsection{User-defined relations and actions}

@@

We are interested in constructing the relational entity graph in Figure~\ref{fig:intromotiv}.
Let $e_1,o_1,r,o_2,$ and $e_2$ be identifiers to the sub-expressions 
$(P|N)+$,$O?$,$R$,$O\wedge 2$, and $(P|N|U)+$, respectively.
%We select $(P|N)+$ to denote an entity as it captures a composite place name such as \cci{Khalifa tower}.
We define the semantic relations $\langle e_1, e_2, r\rangle$, $\langle r, e_1, o_1\rangle$, and $\langle r,e_2,o_2\rangle$.
The matches to $e_1$, $r$, $o_2$, and $e_2$ in match 1 (Fig.~\ref{fig:taskMRE}) are \RL{brj _hlyfT} (Khalifa Tower), \RL{bAlqrb} (next), \RL{mn} (to), and \RL{AltqA.t` Al-'wl} (first intersection).
Note that there is no match to the optional $O$ formula in match 1.
Similarly, the matches to $e_1$, $o_1$, $r$, $o_2$, and $e_2$ in the second match tree are
\RL{dby mwl} (Dubai Mall), \RL{`l_A} (is located), \RL{mqrbT} (near), \RL{mn h_dA} (this), and \RL{Almbn_A} (building), respectively.

Relation $\langle e_1,e_2,r\rangle$ creates the 
edge labeled \cci{next to} between \cci{Khalifa tower} and 
\cci{intersection 1} nodes from match 1, and the 
edge labeled \cci{near} between \cci{Dubai Mall} and \cci{the building} nodes from match 2.
Relation $\langle r,e_1,o_1\rangle$ creates the edge labeled \cci{prep} 
between \cci{Dubai Mall} and \cci{near} nodes from match 2. 
Relation $\langle r,e_2,o_2\rangle$ creates the edge labeled \cci{from} 
between \cci{intersection 1} and \cci{next to} nodes in 
match 1, and the 
edge labeled \cci{from this} between \cci{near} and \cci{the building} 
nodes in match 2.

After the relation construction phase, we are interested to relate the discovered entities
% and relational entities 
that express the same concept.
\framework provides the extended synonym feature of second order as a default cross-reference relation ($Syn^2$).
%\framework provides the \cci{isA} predicate as a default cross-reference relation.
In Figure~\ref{fig:intromotiv}, triggering this feature creates the edge between the nodes \cci{Khalifa Tower} and \cci{The building}.
The edge is labeled with \cci{isA}.

Moreover, \framework allows advanced users to write C++ code snippets 
to process matches of sub-expressions.
Each sub-expression can be associated with two computational actions: \cci{pre-match} and \cci{on-match}.
Users can use these actions to compute statistical features, 
store intermediate results, 
or apply intelligent entity inference techniques
as we show later in the numerical extraction example of Subsection~\ref{subsec:numnorm}.
\framework provides an API that enriches the actions with detailed access to
all solution features of an expression or a formula match including 
text, position, length, equivalent numerical value when applicable, 
and morphological features.
Once \framework computes all match trees, it traverses each tree to 
execute the user defined \cci{pre-match} actions in pre-order manner
and the \cci{on-match} actions in post-order manner.

\begin{comment}
\subsection{\framework simulator}

@@

%\subsection{MRE and action simulation}

For each MRE, 
\framework generates its equivalent non-deterministic finite automaton (NFA) in the typical manner~\cite{sipser2006introduction}. 
%Each MRE operation has its equivalent representation in an NFA. 
We support the upto operation ($f$\^{}$x$), which is not directly 
supported in ~\cite{sipser2006introduction}, by 
%it is equivalent to $f?|ff|\dots|\underbrace{f}_{x \text{ times}}$ which has an NFA mapping. 
%we can expand it into a standard regular expression form, for example
expanding it into a regular expression form; for example 
$f$\^{}$3$ is equivalent to $f?|ff|fff$. 
Consider the example of Figure~\ref{fig:intromotiv} and the
corresponding expression $(P|N)\!+~O?~R~O^\wedge 2~(P|N|U)+$. 
Figure~\ref{fig:nfaEx} shows part of the corresponding NFA where
$q_8, q_9, \dots, q_{13}$ represent NFA states,
and edges are transitions based on MBF tags such as 
$P,$ and $N$.
Edges labeled with the empty string $\epsilon$ are non-deterministic.
%The expression uses operations $|$, $?$, $+$, and $\wedge$ to relate places {\tt P}, names of persons {\tt N}, relative positions {\tt R}, numerical terms {\tt U}, and other {\tt O}.

\framework simulates the generated NFA over the sequence of tag sets matching the MBF formulae.
A simulation match $m$ of an expression $f$ is a tree where the root is the 
MRE expression, the internal nodes are the MRE and MBF operations, and the 
leaves are matches of the MAT terms of $f$.
The leave matches form a vector of tags
$\langle r_k,r_{k+1},\dots,r_j\rangle$ 
corresponding to the text sequence 
$\langle t_k,t_{k+1},\dots,t_j\rangle$ where $r_{\ell}\in R_{\ell},0\le k\le \ell \le j < n$.
%
%If the simulation has a match $\langle r_m,r_{m+1},\dots,r_n\rangle$ 
%where $0\le m\le n$, 
%the next simulation starts at $R_{n+1}$. 
%This disallows overlap of matches for the same MRE. 
%
%In case the NFA simulation has no match, 
%the next simulation starts at $R_{m+1}$. 
If we have more than one match %starting at $R_k$ where $0\le k\le n$, 
\framework returns the longest. 
%
Figure~\ref{fig:motiv}(b) shows two match trees of $\mathit{dir}$ extracted 
from the text of Figure~\ref{fig:intromotiv}.
\RL{dby} and \RL{mwl} are leaf nodes referring to name of place tags ($P$). 
The $+$, $|$, and $?$ MRE operations
are internal nodes.

@@

%\begin{figure}[tb!]
%{ \relsize{-1}
%\begin{framed}
%\begin{multicols}{2}
%\begin{itemize}
%\item Annotated Expression: \\
%\\
%{ \relsize{-2.5}
%%\begin{itemize}
%%  \item 
%  $\stackrel{e_1}{(P|N)+}~\stackrel{o_1}{\mathit{O}?}~\stackrel{r}{R}~\stackrel{o_2}{\mathit{O^{\wedge}2}}~\stackrel{e_2}{(P|N|U)+}$
%%\end{itemize}
%}
%\end{itemize}
%\begin{itemize}
%\item User defined semantic \\relations:
%\begin{itemize}
%\item $\langle e_1,e_2,r\rangle$
%\item $\langle r,e_1,o_1\rangle$
%\item $\langle r,e_2,o_2\rangle$
%\end{itemize}
%\end{itemize}
%\columnbreak
%\setcode{utf8}
%\transfalse
%\resizebox{0.9\columnwidth}{!}{
%	\relsize{+1.5} \input{figures/ergraphEx.tex}
%}
%\transtrue
%\setcode{standard}
%\end{multicols}
%\end{framed}
%}
%\caption{User-defined semantic relation example}
%\label{fig:srEx}
%\end{figure}

\framework computes the relational entities in a user defined 
relation $R=\ldbrack \langle e_1,e_2,r\rangle \rdbrack$
$\subseteq \ldbrack e_1 \rdbrack \times \ldbrack e_2 \rdbrack \times \ldbrack r \rdbrack$ 
to be the elements of $\ldbrack e_1 \rdbrack \times \ldbrack e_2 \rdbrack \times \ldbrack r \rdbrack$
with the smallest nonzero positive distance between the source and the destination
where the distance is the number of words between the matches.

%\framework supports the following configurations for the subexpressions of $e_i$ and $e_j$:
%\begin{itemize}
%\item $\left\vert{e_i}\right\vert=1\wedge\left\vert{e_j}\right\vert=1$: Relate $e_i$ match to $e_j$ match with edge labelled by $r$.
%\item $i=j\wedge\left\vert{e_i}\right\vert>1$: Relate sequences of $e_i$ matches with edges labelled by $r$.
%\item $i\neq j\wedge\left\vert{e_i}\right\vert>1\wedge\left\vert{e_j}\right\vert=1$: Relate each $e_i$ match to the single $e_j$ match with an edge labelled by $r$. The reverse applies for $\left\vert{e_i}\right\vert=1\wedge\left\vert{e_j}\right\vert>1$.
%\item $i\neq j\wedge\left\vert{e_i}\right\vert>1\wedge\left\vert{e_j}\right\vert>1$: Relate each $e_i$ match to each $e_j$ match with an edge labelled by $r$.
%\end{itemize}
%
In Figure~\ref{fig:motiv}(b), \framework names the subexpressions 
$(P|N)+$, $(P|N|U)+$, $O?$, $O\wedge 2$, and $R$, 
as $e_1, e_2, o_1, o_2,$ and $r$, respectively. 
The user defines the semantic relations
$\langle e_1, e_2, r\rangle$, 
$\langle r, e_1, o_1\rangle$, and 
$\langle r,e_2,o_2\rangle$.

The matches of $e_1,~e_2,~o_1,~o_2,$ and $r$ from the second match tree 
of Figure~\ref{fig:motiv}(b) are \RL{dby mwl}(Dubai Mall), 
\RL{Almbn_A}(the building), \RL{`l_A}(prep), \RL{mn h_dA}(from this), 
and \RL{mqrbT}(near), respectively. 
\framework~constructs the semantic relation matches and builds the lower part of the 
entity-relation graph shown in Figure~\ref{fig:intromotiv}.

For the first match \RL{brj _hlyfT bAlqrb mn AltqA.t` Al-'-wl},
the matches of $e_1,~e_2,~o_2,$ and $r$ are 
\RL{brj _hlyfT}(Khalifa tower), \RL{AltqA.t` Al-'-wl}(intersection 1), 
\RL{mn}(from), and \RL{bAlqrb}(next to), respectively. 
\framework~doesn't construct the relation $\langle r, e_1, o_1\rangle$ 
since $o_1$ has no match. 
Therefore, we get the upper part of the entity-relation graph shown 
in Figure~\ref{fig:intromotiv}.

After computing the relational entities, 
\framework~computes a cross-reference relation between the extracted entities
using a second order synonymy feature ($Syn^2$).
The \cci{isA} edge in the 
graph of Figure~\ref{fig:intromotiv} shows the cross-reference relation 
between \RL{brj _hlyfT}(Khalifa tower) from the first match 
with \RL{Almbn_A}(the building) from the second match.
\end{comment}

>>>>

@@

\major{
Let ${\cal O} = \{ \mathit{isA}, \mathit{contains}\}$ be the set of atomic term 
predicates, where $\mathit{isA}$ and $\mathit{contains}$ denote exact match and containment, respectively.
Also, let ${\cal F} = \{ \pp, \ss, \xx, \PP, \GG, \AC\}$ be the 
set of morphological features where each morphological feature $A\in {\cal F}$ is in 
turn a set of morphological feature values.}
Given a word $w$,
a user defined constant feature value $CF\in A$,
and an integer $k, 1\le k\le 7$, 
the following are morphology-based atomic terms (MAT), {\em terms} for short.
\begin{itemize}
  \item $a(w):= \exists m \in M(w).~m=\langle p,s,x,P,G,C\rangle. r \circ CF$
where $\circ \in {\cal O}$, $r \in \{p,s,x,P,G,C\}$, and $r\in A$.
Informally, a solution vector of $w$ exists with
a feature containing or exactly matching \major{the user-chosen feature value} $CF$.
\item $a(w) := w \in Syn^k(CF), CF \in \ss$.
  Informally, this checks if $w$ is an extended synonym of a stem $CF$. 
\major{
  We limit $k$ to a maximum of $7$ since we practically noticed that 
  (1) values above $7$ introduce significant semantic noise and
  (2) the computation is expensive without a bound. 
}
\end{itemize}

@@

\major{
Consider the task we discussed in the introduction (Figure~\ref{fig:intromotiv}) 
and recall that we are interested in identifying names of people, names of places, relative positions, and numerical terms.
The table in Figure~\ref{fig:taskMBF} presents the defined formulae.
The user denotes the ``name of person'' entities with formula $N$ 
which requires the {\em category} feature in the morphological solution of a word to be {\tt Name\_of \_Person}.
The entities $n_1$, $n_2$, and $n_3$ are matches of the formula $N$ in the text.
Similarly, the user specifies formula $P$ to denote ``name of place'' entities. 
The user specifies formula $R$ to denote ``relative position'' entities, 
and defines it as a disjunction of terms that check for 
solutions matching stems such as \RL{qrb} (near) and \RL{fy} (in).
%requires the stem feature of a word to belong to a selected list of stems containing \RL{fy} and \RL{qrb}. a
Similarly, $U$ denotes numerical terms and is a disjunction of constraints 
requiring the stem feature to belong to a set of stems such as
\RL{'wl}(first), \RL{_tAny}(second), \dots \RL{`A^sr}(tenth).
}

@@

Next, we define a morphology-based regular expression (MRE) as follows.
\begin{itemize}
\item $m$ is an MRE where $m$ is an MBF.
\item $fg$ is an MRE 
  \major{
  where $f$ and $g$ are both MRE expressions. 
    A match of $f$ followed by a match of $g$ satisfies this concatenation operation.}
\item $f*,f+,f$\textasciicircum$x,$ and $f?$ \major{ are MRE } where $f$ is an MRE,
  and are satisfied by zero or more, \major{ one or }more, up to $x$ matches, 
    and an \major{optional single match} of $f$, respectively.
  \item $f\& g,$ (conjunction) and $f|g$ (disjunction) are \major{MRE} where 
    $f$ and $g$ are \major{MRE}, and are satisfied by 
    \major{the intersection of $f$ and $g$ matches, 
    and the union of the $f$ and $g$ matches}, respectively. 
\end{itemize}
We denote by $\ldbrack f\rdbrack$ the set of matches of an MRE $f$.

\major{
Back to the example in Figure~\ref{fig:intromotiv}.
We use the formulae defined in Figure~\ref{fig:taskMBF} to construct an MRE such as
$(P|N)\!+O?~R~O^\wedge\!2~(P|N|U)+$ where 
$|,+,?,$ and $^\wedge k$ denote disjunction, one or more, zero or one, and
up to $k$ matches, respectively.
The expression specifies a sequence of places or names of persons, 
optionally followed by a null word, 
followed by one relative position, followed by up to two possible null words, 
followed by one or more match of name of place, name of person, or numerical term.
$O?$ and $O^\wedge 2$ are used in the expression to allow for flexible matches.
%Note that the user reaches the satisfying matches by experimenting with the visualizer and the expression editor which do not require knowledge and expertise in regular expressions. 
}

@@

\major{
The \major{matching parse trees} in Figure~\ref{fig:taskMRE} illustrate two matches of the expression computed by \framework.
The first tree refers to the text 
\RL{brj _hlyfT bAlqrb mn AltqA.t` Al-'wl}(Khalifa Tower next to the first intersection).
The second tree refers to the text 
\RL{dby mwl `l_A mqrbT mn h_dA Almbn_A}(Dubai Mall is located near this building).
}
%The leaf nodes of the trees are entities and the edges and internal nodes are text, 
%morphology-based, 
%and word distance based relational entities. 
\major{
The leaf nodes of the trees are matches to formulae and the internal nodes represent roots to sub-expression matches.
For instance, \RL{brj _hlyfT} in match 1 tree corresponds to the sub-expression $(P|N)+$.
}

\subsection{User-defined relations and actions}
\label{sec:relations-actions}

@@

\major{
We are interested in constructing the relational entity graph in Figure~\ref{fig:intromotiv}.
Let $e_1,o_1,r,o_2,$ and $e_2$ be identifiers to the sub-expressions 
$(P|N)+$,$O?$,$R$,$O\wedge 2$, and $(P|N|U)+$, respectively.
%We select $(P|N)+$ to denote an entity as it captures a composite place name such as \cci{Khalifa tower}.
We define the semantic relations $\langle e_1, e_2, r\rangle$, $\langle r, e_1, o_1\rangle$, and $\langle r,e_2,o_2\rangle$.
The matches to $e_1$, $r$, $o_2$, and $e_2$ in match 1 (Fig.~\ref{fig:taskMRE}) are \RL{brj _hlyfT} (Khalifa Tower), \RL{bAlqrb} (next), \RL{mn} (to), and \RL{AltqA.t` Al-'wl} (first intersection).
Note that there is no match to the optional $O$ formula in match 1.
Similarly, the matches to $e_1$, $o_1$, $r$, $o_2$, and $e_2$ in the second \major{matching parse tree} are
\RL{dby mwl} (Dubai Mall), \RL{`l_A} (is located), \RL{mqrbT} (near), \RL{mn h_dA} (this), and \RL{Almbn_A} (building), respectively.
}

\major{
Relation $\langle e_1,e_2,r\rangle$ creates the 
edge labeled \cci{next to} between \cci{Khalifa tower} and 
\cci{intersection 1} nodes from match 1, and the 
edge labeled \cci{near} between \cci{Dubai Mall} and \cci{the building} nodes from match 2.
Relation $\langle r,e_1,o_1\rangle$ creates the edge labeled \cci{prep} 
between \cci{Dubai Mall} and \cci{near} nodes from match 2. 
Relation $\langle r,e_2,o_2\rangle$ creates the edge labeled \cci{from} 
between \cci{intersection 1} and \cci{next to} nodes in 
match 1, and the 
edge labeled \cci{from this} between \cci{near} and \cci{the building} 
nodes in match 2.
}

Moreover, \framework allows advanced users to write C++ code snippets 
to process matches of sub-expressions.
Each sub-expression can be associated with two computational 
actions: \cci{pre-match} and \cci{on-match}.
%
\framework provides an API that enriches the actions with detailed access to
all solution features of an expression or a formula match including 
text, position, length, equivalent numerical value when applicable, 
and morphological features.
\major{
The API follows a decorator pattern in that it incrementally adds the 
action results to the matching entities. 
}
Once \framework computes all \major{matching parse trees}, it traverses each tree to 
execute the user defined \cci{pre-match} actions in pre-order manner
and the \cci{on-match} actions in post-order manner.
\major{
This follows an observer pattern that notifies listeners with each
produced match. 
}

\subsection{\framework simulators}

@@

For each MRE, 
\framework generates its equivalent non-deterministic finite automaton (NFA) in the typical manner~\cite{sipser2006introduction}.
%Each MRE operation has its equivalent representation in an NFA. 
We support the upto operation ($f$\^{}$x$), which is not directly 
supported in~\cite{sipser2006introduction}, by 
%it is equivalent to $f?|ff|\dots|\underbrace{f}_{x \text{ times}}$ which has an NFA mapping. 
%we can expand it into a standard regular expression form, for example
expanding it into a regular expression form; for example 
$f$\^{}$3$ is equivalent to $f?|ff|fff$. 
Consider the example of Figure~\ref{fig:intromotiv} and the
corresponding expression $(P|N)\!+~O?~R~O^\wedge 2~(P|N|U)+$. 
Figure~\ref{fig:nfaEx} shows part of the corresponding NFA where
$q_8, q_9, \dots, q_{13}$ represent NFA states,
and edges are transitions based on MBF tags such as 
$P,$ and $N$.
Edges labeled with the empty string $\epsilon$ are non-deterministic.
%The expression uses operations $|$, $?$, $+$, and $\wedge$ to relate places {\tt P}, names of persons {\tt N}, relative positions {\tt R}, numerical terms {\tt U}, and other {\tt O}.

\framework simulates the generated NFA over the sequence of tag sets matching the MBF formulae.
%A simulation match $m$ of an expression $f$ is a tree where the root is the MRE expression, the internal nodes are the MRE and MBF operations, and the leaves are matches of the MAT terms of $f$.
A simulation match $m$ of an expression $f$ is a tree where the root spans the expression, the internal nodes are roots to sub-expressions of $f$, and the leaves are matches of the MBF formulae of $f$, e.g. Fig.~\ref{fig:taskMRE}.
The sequence of leaf matches forms a vector of tags $\langle r_k,r_{k+1},\dots,r_j\rangle$ 
corresponding to the text sequence 
$\langle t_k,t_{k+1},\dots,t_j\rangle$ where $r_{\ell}\in R_{\ell},0\le k\le \ell \le j < n$.
%
%If the simulation has a match $\langle r_m,r_{m+1},\dots,r_n\rangle$ 
%where $0\le m\le n$, 
%the next simulation starts at $R_{n+1}$. 
%This disallows overlap of matches for the same MRE. 
%
%In case the NFA simulation has no match, 
%the next simulation starts at $R_{m+1}$. 
If we have more than one match for an expression, 
%starting at $R_k$ where $0\le k\le n$, 
\framework returns the longest.
%Figure~\ref{fig:taskMRE} shows two match trees of $\mathit{dir}$ extracted from the text of Figure~\ref{fig:intromotiv}.
%\RL{dby} and \RL{mwl} are leaf nodes referring to name of place tags ($P$).
%The $+$, $|$, and $?$ MRE operations are internal nodes.

@@

%\begin{figure}[tb!]
%{ \relsize{-1}
%\begin{framed}
%\begin{multicols}{2}
%\begin{itemize}
%\item Annotated Expression: \\
%\\
%{ \relsize{-2.5}
%%\begin{itemize}
%%  \item 
%  $\stackrel{e_1}{(P|N)+}~\stackrel{o_1}{\mathit{O}?}~\stackrel{r}{R}~\stackrel{o_2}{\mathit{O^{\wedge}2}}~\stackrel{e_2}{(P|N|U)+}$
%%\end{itemize}
%}
%\end{itemize}
%\begin{itemize}
%\item User defined semantic \\relations:
%\begin{itemize}
%\item $\langle e_1,e_2,r\rangle$
%\item $\langle r,e_1,o_1\rangle$
%\item $\langle r,e_2,o_2\rangle$
%\end{itemize}
%\end{itemize}
%\columnbreak
%\setcode{utf8}
%\transfalse
%\resizebox{0.9\columnwidth}{!}{
%	\relsize{+1.5} \input{figures/ergraphEx.tex}
%}
%\transtrue
%\setcode{standard}
%\end{multicols}
%\end{framed}
%}
%\caption{User-defined semantic relation example}
%\label{fig:srEx}
%\end{figure}
%
Next, \framework computes the relational entities corresponding to each user defined relation $\ldbrack \langle e_1,e_2,r\rangle \rdbrack$
$\subseteq \ldbrack e_1 \rdbrack \times \ldbrack e_2 \rdbrack \times \ldbrack r \rdbrack$.

\begin{comment} 
Note that the number of matches for $e_i$ ($\ldbrack e_i \rdbrack$) depends on the corresponding sub-expression.
For instance, we could have multiple matches if we associate $e_i$ with the sub-expression $P|N$ while we have $(P|N)+$.
%to be the elements of $\ldbrack e_1 \rdbrack \times \ldbrack e_2 \rdbrack \times \ldbrack r \rdbrack$
%with the smallest nonzero positive distance between the source and the destination
%where the distance is the number of words between the matches. 
Hence, \framework supports the following configurations for $e_i$ and $e_j$ based on the corresponding matches of sub-expressions:
\begin{itemize}
\item $\left\vert{e_i}\right\vert=1\wedge\left\vert{e_j}\right\vert=1$: Relate $e_i$ match to $e_j$ match with edge labeled by $r$.
\item $i=j\wedge\left\vert{e_i}\right\vert>1$: Relate sequences of $e_i$ matches with edges labeled by $r$.
\item $i\neq j\wedge\left\vert{e_i}\right\vert>1\wedge\left\vert{e_j}\right\vert=1$: Relate each $e_i$ match to the single $e_j$ match with an edge labeled by $r$. The reverse applies for $\left\vert{e_i}\right\vert=1\wedge\left\vert{e_j}\right\vert>1$.
\item $i\neq j\wedge\left\vert{e_i}\right\vert>1\wedge\left\vert{e_j}\right\vert>1$: Relate each $e_i$ match to each $e_j$ match with an edge labeled by $r$.
\end{itemize}
\end{comment}

%In Figure~\ref{fig:taskMRE}, \framework names the sub-expressions $(P|N)+$, $(P|N|U)+$, $O?$, $O\wedge 2$, and $R$, as $e_1, e_2, o_1, o_2,$ and $r$, respectively. 
%The user defines the semantic relations $\langle e_1, e_2, r\rangle$, $\langle r, e_1, o_1\rangle$, and $\langle r,e_2,o_2\rangle$.

%The matches of $e_1,~e_2,~o_1,~o_2,$ and $r$ from the second match tree of Figure~\ref{fig:motiv}(b) are \RL{dby mwl}(Dubai Mall), \RL{Almbn_A}(the building), \RL{`l_A}(prep), \RL{mn h_dA}(from this), and \RL{mqrbT}(near), respectively. 
%\framework~constructs the semantic relation matches and builds the lower part of the entity-relation graph shown in Figure~\ref{fig:intromotiv}.

%For the first match \RL{brj _hlyfT bAlqrb mn AltqA.t` Al-'-wl}, the matches of $e_1,~e_2,~o_2,$ and $r$ are \RL{brj _hlyfT}(Khalifa tower), \RL{AltqA.t` Al-'-wl}(intersection 1), \RL{mn}(from), and \RL{bAlqrb}(next to), respectively. 
%\framework doesn't construct the relation $\langle r, e_1, o_1\rangle$ since $o_1$ has no match. 
%Therefore, we get the upper part of the entity-relation graph shown in Figure~\ref{fig:intromotiv}.

%The \cci{isA} edge in the graph of Figure~\ref{fig:intromotiv} shows the cross-reference relation between \RL{brj _hlyfT}(Khalifa tower) from the first match with \RL{Almbn_A}(the building) from the second match.
